---
title: "SDS 315 Homework 8 Response"
author: "Grace Lefferts (gel766)"
date: "2025-04-14"
output: pdf_document
urlcolor: blue
---

[GitHub Link](https://github.com/graceleff05/SDS315_HW8)

```{r global_options, echo=FALSE}
knitr::opts_chunk$set(warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80), echo=FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(kableExtra)
library(mosaic)
library(dplyr)
creatinine = read.csv('creatinine.csv')
covid = read.csv('covid.csv')
milk = read.csv('milk.csv')
```

## **Problem 1: Regression Warm Up**

```{r, results = 'hide'}
# Fitting regression model to creatinine data
lm_creat = lm(creatclear ~ age, data = creatinine)
coef(lm_creat)
```

**A)** We should expect a creatinine clearance rate of 113.71 mL/minute for a 55-year-old. I determined this clearance rate by first fitting a regression model to the creatinine data to predict creatinine clearance rate based on age, which results in the equation y = 147.81 - 0.62x, where y is predicted creatinine clearance rate in mL/min and x is age in years. Secondly, to calculate the expected clearance rate for a 55 y/o, I simply plugged 55 into x and solved for y.

**B)** Creatinine clearance rate decreases at a rate of -0.62 mL/minute per year. I determined this based on the slope of the regression model I calculated in the previous question.

**C)** A 40-year-old with a creatinine clearance rate of 135 is healthier for their age than a 60-year-old with a rate of 112. I determined this based on their respective residuals, which I calculated by subtracting their actual rates from their expected rates (which I calculated using the regression model). Because the residual for the 40-year-old is higher than that of the 60-year-old (11.99 vs. 1.39), the 40-year-old is considered healthier for their age.

-   Expected rate for 40 y/o: 147.81 - 0.62\*40 = 123.01
-   Residual for 40 y/o: 135 - 123.01 = 11.99
-   Expected rate for 60 y/o: 147.81 - 0.62\*60 = 110.61
-   Residual for 60 y/o: 112 - 110.61 = 1.39


## **Problem 2: Modeling Disease Growth**

### *Italy*
```{r, results = 'hide'}
# Subsetting covid data to include only Italy
covid_Italy = covid %>%
  filter(country == 'Italy')

# Fitting an exponential growth model
lm_Italy = lm(log(deaths) ~ days_since_first_death, data=covid_Italy)
coef(lm_Italy)

# 95% bootstrapped confidence interval for growth rate
boot_covid_Italy = do(10000)*lm(log(deaths) ~ days_since_first_death, data=mosaic::resample(covid_Italy))
confint(boot_covid_Italy, level = 0.95) %>% filter(name == 'days_since_first_death')

# Calculating doubling time
DT_Italy = round(70/18.3, 1)

# 95% bootstrapped confidence interval for doubling time
upperItaly = round(70/15.9, 1)
lowerItaly = round(70/20.8, 1)
```

-   Estimated growth rate: 0.183 (or 18.3%) daily growth rate in deaths
-   Estimated growth rate CI: [0.159, 0.208]
-   Doubling time: 3.8 days
-   Doubling time CI: [3.4, 4.4]

### *Spain*
```{r, results = 'hide'}
# Subsetting covid data to include only Spain
covid_Spain = covid %>%
  filter(country == 'Spain')

# Fitting an exponential growth model
lm_Spain = lm(log(deaths) ~ days_since_first_death, data=covid_Spain)
coef(lm_Spain)

# 95% bootstrapped confidence interval for growth rate
boot_covid_Spain = do(10000)*lm(log(deaths) ~ days_since_first_death, data=mosaic::resample(covid_Spain))
confint(boot_covid_Spain, level = 0.95) %>% filter(name == 'days_since_first_death')

# Calculating doubling time
DT_Spain = round(70/27.6, 1)

# 95% bootstrapped confidence interval for doubling time
upperSpain= round(70/23.5, 1)
lowerSpain = round(70/31.8, 1)
```

-   Estimated growth rate: 0.276 (or 27.6%) daily growth rate in deaths
-   Estimated growth rate CI: [0.235, 0.318]
-   Doubling time: 2.5 days
-   Doubling time CI: [2.2, 3.0]

### *Daily Deaths Over Time*
```{r}
ggplot(covid) +
  geom_line(aes(x = days_since_first_death, y = deaths, color = country)) +
  labs(title= 'Reported Daily Deaths Over Time by Country') +
  theme_minimal()
```


## **Problem 3: Price Elasticity of Demand**

```{r, results = 'hide'}
# Fitting a power law model 
lm_milk = lm(log(sales) ~ log(price), data=milk)
coef(lm_milk)

# 95% bootstrapped confidence interval for elasticity
boot_milk = do(10000)*lm(log(sales) ~ log(price), data=mosaic::resample(milk))
confint(boot_milk, level = 0.95) %>% filter(name == 'log.price.')
```
The estimated price elasticity of the demand for milk is -1.62 (meaning that for every 1% increase in the price of milk, the demand decreases by 1.62%). I determined this by fitting a power law model to the milk data set (calculating a linear model with x = log(price) and y = log(sales)) and identifying the coefficient for log(price). Based on the bootstrapped 95% confidence interval I calculated, the price elasticity is somewhere between -1.77 and -1.45.

